---
title: "Create mock data to test DrugUtilisation package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DrugUtilisation mock data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE}
library(DrugUtilisation)
library(dplyr, warn.conflicts = FALSE)
```

# Introduction

In this vignette we will see how to use `mockDrugUtilisation()` function to create mock data. This function is predominantly used in this package's unit testing.

For example, one could use the default parameters to create a mock cdm reference like so:

```{r, message= FALSE, warning=FALSE}
cdm <- mockDrugUtilisation()
```

As you can see this creates a local (in memory) cdm reference object:

```{r}
cdm
```

This will then populate several omop tables (for example, `person`, `concept` and `visit_occurrence`) and two cohorts in the cdm reference.

```{r, message= FALSE, warning=FALSE}
cdm$person |>
  glimpse()

cdm$person |>
  tally()

cdm$concept |>
  glimpse()

cdm$concept |>
  tally()

cdm$visit_occurrence |>
  glimpse()

cdm$visit_occurrence |>
  tally()

cdm$cohort1 |>
  glimpse()

cdm$cohort1 |>
  tally()

cdm$cohort2 |>
  glimpse()

cdm$cohort2 |>
  tally()
```

## Insert to `duckdb`

By default as we have seen the generated mock data is locally stored in memory which is not a realistic situation in some cases. So inserting this cdm_reference to a database is quite useful in terms to simulate a real situation. You can insert the local cdm to another source (e.g. duckdb, postgres, sql server, arrow...) using the function `omopgenerics::insertCdmTo()`. Alternatively you can use the `source` argument to insert the cdm to a *DuckDB* database:

```{r}
cdm <- mockDrugUtilisation(source = "duckdb")
cdm
```

## Setting seeds

The user can also set the seed to control the randomness within the data.

```{r, message= FALSE, warning=FALSE}
set.seed(seed = 1)
cdm <- mockDrugUtilisation(source = "duckdb")
cdm$person |>
  glimpse()
```

If we would run it again the result will be different:

```{r}
cdm <- mockDrugUtilisation(source = "duckdb")
cdm$person |>
  glimpse()
```

But if we set the same seed again, the result would be the same than initially:

```{r}
set.seed(seed = 1)
cdm <- mockDrugUtilisation(source = "duckdb")
cdm$person |>
  glimpse()
```

This can be quite useful to generate the same cdm_reference object.

## Create bigger mock

By default the generated mock contains only 10 individuals and small tables size:

```{r}
lapply(cdm, \(x) x |> tally() |> pull())
```

You can change that using the `numberIndividual` argument to generate more individuals and records:

```{r, message= FALSE, warning=FALSE}
cdm <- mockDrugUtilisation(numberIndividual = 100, source = "duckdb")
```

This will ensure that the `person` table will define 100 mock individuals: 

```{r, message= FALSE, warning=FALSE}
cdm$person |>
  tally()
```

As a consequence of this, the generated tables will have more rows compared to the initial ones:

```{r, message= FALSE, warning=FALSE}
lapply(cdm, \(x) x |> tally() |> pull())
```

## Creat mock data by customising tables

### Customise omop tables

As we saw previously, the omop tables are automatically populated in `mockDrugUtilisation()`. However, the user can customise these tables. For example, to customise `drug_exposure` table, one could do the following:

```{r, message= FALSE, warning=FALSE}
cdm <- mockDrugUtilisation(
  drug_exposure = tibble(
    drug_exposure_id = 1:3,
    person_id = c(1, 1, 1),
    drug_concept_id = c(2, 3, 4),
    drug_exposure_start_date = as.Date(c(
      "2000-01-01", "2000-01-10", "2000-02-20"
    )),
    drug_exposure_end_date = as.Date(c(
      "2000-02-10", "2000-03-01", "2000-02-20"
    )),
    quantity = c(41, 52, 1),
    drug_type_concept_id = 0
  ),
  source = "duckdb"
)
```

```{r, message= FALSE, warning=FALSE}
cdm$drug_exposure |>
  glimpse()
```

However, one needs to be vigilant that the customised omop table is implicitly dependent on other omop tables. For example, the `drug_exposure_start_date` of someone in the `drug_exposure` table should lie within that person's `observation_period_start_date` and `observation_period_end_date`.

One could also modify other omop tables including `person`, `concept`, `concept_ancestor`, `drug_strength`, `observation_period`, `condition_occurrence`, `observation`, and `concept_relationship` in a similar fashion.

### Customise cohorts

Similarly, cohort tables can also be customised.

```{r, message= FALSE, warning=FALSE}
cdm <- mockDrugUtilisation(
  observation_period = tibble(
    observation_period_id = 1,
    person_id = 1:2,
    observation_period_start_date = as.Date("1900-01-01"),
    observation_period_end_date = as.Date("2100-01-01"),
    period_type_concept_id = 0
  ),
  cohort1 = tibble(
    cohort_definition_id = 1,
    subject_id = c(1, 1, 2),
    cohort_start_date = as.Date(c("2000-01-01", "2001-01-01", "2000-01-01")),
    cohort_end_date = as.Date(c("2000-03-01", "2001-03-01", "2000-03-01"))
  ),
  source = "duckdb"
)
```

```{r, message= FALSE, warning=FALSE}
cdm$cohort1 |>
  glimpse()
```
