---
title: "Summarise treatments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarise treatments}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

After creating a study cohort, for example of some specific condition of interest, we may be interested in describing the treatments received by the individuals within it. Here we show how such a summary can be obtained.

## Create mock table

We will use mock data contained in the package throughout the vignette. Let's
modify cohort tables `cohort1` and `cohort2` in our mock dataset, so the first table includes 3 cohorts of health conditions (our study cohorts), while the second contains three are of treatments they could receive.

```{r, message = FALSE, warning=FALSE}
library(DrugUtilisation)

cdm <- mockDrugUtilisation(numberIndividual = 200)

new_cohort_set <- settings(cdm$cohort1) |>
  dplyr::arrange(cohort_definition_id) |>
  dplyr::mutate(cohort_name = c("asthma", "bronchitis", "pneumonia"))

cdm$cohort1 <- cdm$cohort1 |>
  omopgenerics::newCohortTable(cohortSetRef = new_cohort_set)

new_cohort_set <- settings(cdm$cohort2) |>
  dplyr::arrange(cohort_definition_id) |>
  dplyr::mutate(cohort_name = c("albuterol", "fluticasone", "montelukast"))

cdm$cohort2 <- cdm$cohort2 |>
  omopgenerics::newCohortTable(cohortSetRef = new_cohort_set)
```

Notice that `cohort1` is a cohort table with three cohorts representing
three different conditions:

```{r, message = FALSE, warning=FALSE}
settings(cdm$cohort1)
```

And `cohort2` is a cohort table with three different treatment cohorts:

```{r, message = FALSE, warning=FALSE}
settings(cdm$cohort2)
```

# Summarise treatment 

The `summariseTreatment()` function produces a summary of the treatment received by our study cohorts. There are three mandatory arguments:

1.  `cohort`: cohort from the cdm object.
2.  `treatmentCohortName`: name of the treatment cohort table.
3.  `window`: a list specifying the time windows during which treatments should be summarised.

See an example of its usage below, where we use
`summariseTreatment()` to summarise treatments defined in
`cohort2` in the target cohorts defined in `cohort1`.

```{r, message = FALSE, warning=FALSE}
summariseTreatment(
  cohort = cdm$cohort1,
  treatmentCohortName = c("cohort2"),
  window = list(c(0, 0), c(1, 30))
)
```

## strata parameter

We can also stratify our cohort and calculate the estimates within each
strata group by using the `strata` parameter.

```{r, message = FALSE, warning=FALSE}
cdm[["cohort1"]] <- cdm[["cohort1"]] |>
  PatientProfiles::addSex() |>
  PatientProfiles::addAge(ageGroup = list("<40" = c(0, 39), ">=40" = c(40, 150)))

results <- summariseTreatment(
  cohort = cdm$cohort1,
  treatmentCohortName = c("cohort2"),
  window = list(c(0, 0)),
  treatmentCohortId = 1,
  strata = list("sex", "age_group")
)
```

Notice that we have also used the `treatmentCohortId` parameter to
specify that we only want to explore `albuterol` (which corresponds to the cohort
id 1 in our cohort table) across the cohorts defined in `cohort1`.

## other parameters

The `summariseTreatment()` functions also has other input parameters which can be tuned:

- `cohortId`: to restrict the analysis to a particular cohort definition id in the target cohort.
- `indexDate`: what column to use as the index date to start the analysis. By default we use `cohort_start_date`, but any other date column, such as `cohort_end_date`, can be specified instead.
- `censorDate`: whether to end the analysis at any specific date. Otherwise we will follow the individuals until end of their respective observation period.
- `mutuallyExclusive`: by default set to FALSE, this will consider the treatments separately, so an individual can belong to different treatment groups at the same time (i.e. if they are treated with multiple drugs). Therefore, for each target cohort, we could have a sum of percentages of all treatment drugs greater than 100%. If set to TRUE, non-overlapping treatment groups will be assessed (with multiple drugs in some of those if needed), so that all individuals will belong to only one of them, and the percentages will add up to a 100.

```{r}
result_not_mutually_exc <- summariseTreatment(
  cohort = cdm$cohort1,
  treatmentCohortName = c("cohort2"),
  window = list(c(0, 0))
)

result_mutually_exc <- summariseTreatment(
  cohort = cdm$cohort1,
  treatmentCohortName = c("cohort2"),
  window = list(c(0, 0)),
  mutuallyExclusive = TRUE
)

tableTreatment(result = result_not_mutually_exc)
tableTreatment(result = result_mutually_exc)
```

In this case, as no individual was given more than one of the treatment drugs, the percentages are the same. However, we can easily see how these analyses would provide different results in other contexts.

# Visualise results

The package includes `table` and `plot` functions to help visualise the results from `summariseTreatment()`, like we have just used to show the results above.

## Tables

The `tableTreatment()` function generates a table in gt, flextable, or tibble format from the summarised_result produced by `summariseTreatment()`. This function has customisation options to format the table according to user preferences.

```{r}
tableTreatment(result = results)
```

## Plots

The `plotTreatment()` function creates a bar plot showing the percentage of treated and untreated in each cohort, stratum, and time-window. This function offers customization options for colors, faceting, and handling of strata.

```{r, fig.height=8, fig.width=8}
plotTreatment(
  result = results,
  facet =  sex + age_group ~ window_name + cohort_name,
  colour = "variable_level"
)
```
